<!--
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%      ___      __         _      
%%     / __\___ / _\  /\/\ (_) ___ 
%%    / /  / _ \\ \  /    \| |/ __|
%%   / /__| (_) |\ \/ /\/\ \ | (__ 
%%   \____/\___/\__/\/    \/_|\___|
%%
%%  COVID-19 Spatial Microsimulation  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Authors:      Christian Dudel
%%               Matthias Rosenbaum-Feldbruegge
%%               Sebastian Kluesener
%%               Ralf Schneider
%%
%% Contact:      dudel@demogr.mpg.de
%%               sebastian.kluesener@bib.bund.de
%%
%%==============================================================================
%%
%% Copyright (C) 2021 Federal Institute for Population Research (BIB),
%%                    The Max Planck Institute for Demographic Research (MPIDR),
%%                    High Performance Computing Center Stuttgart (HLRS)
%% 
%% This program is free software: you can redistribute it and/or modify
%% it under the terms of the GNU General Public License as published by
%% the Free Software Foundation, either version 3 of the License, or
%% (at your option) any later version.
%% 
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%% GNU General Public License for more details.
%% 
%% You should have received a copy of the GNU General Public License
%% along with this program.  If not, see <http://www.gnu.org/licenses/>.
%%==============================================================================
-->
---
title: "CoSMic - Basic usage"
author: "Ralf Schneider <schneider@hlrs.de>"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CoSMic - Basic usage}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction
In this vignette we demonstrate the basic usage of CoSMic function stack to execute **CO**VID-19 **S**patial **Mic**rosimulations to predict coverage of intensive care units.

First things first. Since we want to be better save than sorry we clean the environment.
```{r eval=FALSE}
rm(list=ls())
```
Since CoSMic is currently not developed up to the state of a proper R-package we install it manually from source by means of the devtools and roxygen2 packages.
```{r eval=FALSE}
library(devtools)
library(roxygen2)
devtools::load_all()
devtools::install()
```

After installation we load the package.
```{r eval=FALSE}
library(CoSMic)
```

# Parameter Setup

## Setup of the Execution Control Parameters

### Execution Procedure
First we set the execution procedure.

Valid Values are:

* `Basic-Param` - Multiple model execution and parameter variations.
* `Optimization` - Model parameter optimization by means of the GA package.

```{r}
exec.procedure <- "Basic-Param"
```

### Parallel Execution Method
Valid values :

* `OMP` - Should be used for desktop PCs and shared memory compute nodes.
* `MPI` - Should be used with `Basic-Param` execution procedure on distributed memory architectures.
* `PSOCK` - Should be used with `Optimization` execution procedure on distributed memory architectures to exploit a hirarchical paralleization strategy.

```{r}
parallel.method <- "OMP"
```

### Maximum number of cores
Specify the maximum number of cores to be used on a single node to not run out of memory.

Valid value:

* Any integer
```{r}
max.cores <- 64
```

## Parallel worker debug output
Specify whether std.out from workers should be captured to a file named `cl.out`.

Valid values :

* `TRUE` or `FALSE`

```{r}
omp.cluster.dbg <- TRUE
```

### Path to data directory
Specify the file system path to the directory where the CoSMic input files reside. Which files are essential to provide under wich circumstances will be explained later on.

Valid values:

* Any String valid as a path-name on the used file system
 
```{r}
data.dir <- "Data"
```

### Path to output directory
Specify the file system path to the directory to which the CoSMic result files will be written.

Valid values:

* Any String valid as a path-name on the used file system

```{r}
output.dir <- NULL
```

### Export suffix
To to distinguish different model executions an output suffix can be set.

Valid values are:

* Any String valid to be used in filenames on the used file system.
* `NULL` - Defaults to `v<model_version>-<country>-<exec_date_and_time>`.

```{r}
export_name <- NULL
```

### Write Checkpoint
Specify whether checkpoint data should be written to `output.dir`.

Valid values :

* `TRUE` or `FALSE`

```{r}
cp.write <- FALSE
```

### Timestep of Checkpoint
If `cp.write <- TRUE` at which timestep the checkpoint data should be writtten?

Valid values :

* Any integer smaller than the maximum simulation timestep.

```{r}
cp.time <- 0
```

### Reload checkpoint
Specify whether checkpoint data should be reloaded.

Valid values :

* `TRUE` or `FALSE`
```{r}
cp.reload <- FALSE
```

### Checkpoint Directroy
Specify the path from which checkpoint data should be reloaded in case `cp.reload <- TRUE`

Valid values:

* Any String valid as a path-name on the used file system

```{r}
cp.dir <- NULL
```
-------------------------------------------------

## Setup of the Model Control Parameters

### Seeding 
First we have to specify if and if yes where to seed.

#### Seed Base
Valid values:

* `NULL` for no seeding
* `42  ` for seeding so that Doug will be proud of us!
```{r}
seed.base <- NULL
```

#### Seed within model loop
Specify whether seeding should happen within the model loop. In case of `MPI` or `PSOCK`
used as parallel method this enables deterministic model execution over multiple executions.

Valid values:

* `TRUE` or `FALSE`

If set to `TRUE`,  `seed.base` will also be used in a call to `set.seed` within
the model loop. This will make all iterations of the model parameters equal.
```{r}
seed.in.inner.loop = FALSE
```

### Derivation of initial population

#### Country to simulate

Valid values:

* `"Germany"`

```{r}
country <- c("Germany")
```

#### Population Sample Size
Choose sample size.

Valid values:

* Any integer. E.g.: 8301921 represent a 10% sample of the German population.

In this demonstration we use a rather small sample to keep the required computing
resources within reasonable limits.
```{r warning=FALSE}
CoSMic::set.pspace(param  = "sam_size", values = 83019)
```
As you might have noticed, we use the `set.pspace` function to set this parameter since it
potentially can take part of a parameter space exploration.

