<!--
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%      ___      __         _      
%%     / __\___ / _\  /\/\ (_) ___ 
%%    / /  / _ \\ \  /    \| |/ __|
%%   / /__| (_) |\ \/ /\/\ \ | (__ 
%%   \____/\___/\__/\/    \/_|\___|
%%
%%  COVID-19 Spatial Microsimulation  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Authors:      Christian Dudel
%%               Matthias Rosenbaum-Feldbruegge
%%               Sebastian Kluesener
%%               Ralf Schneider
%%
%% Contact:      dudel@demogr.mpg.de
%%               sebastian.kluesener@bib.bund.de
%%
%%==============================================================================
%%
%% Copyright (C) 2021 Federal Institute for Population Research (BIB),
%%                    The Max Planck Institute for Demographic Research (MPIDR),
%%                    High Performance Computing Center Stuttgart (HLRS)
%% 
%% This program is free software: you can redistribute it and/or modify
%% it under the terms of the GNU General Public License as published by
%% the Free Software Foundation, either version 3 of the License, or
%% (at your option) any later version.
%% 
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%% GNU General Public License for more details.
%% 
%% You should have received a copy of the GNU General Public License
%% along with this program.  If not, see <http://www.gnu.org/licenses/>.
%%==============================================================================
-->
---
title: "CoSMic - Basic usage"
author: "Ralf Schneider <schneider@hlrs.de>"
output:
  html_document:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{CoSMic - Basic usage}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction
In this vignette we demonstrate the usage of the CoSMic function stack to execute
**CO**VID-19 **S**patial **Mic**rosimulations to predict coverage of intensive care units.

At this place we demonstrate the execution of CoSMic with minimal paramter setup efforts
so keeping basically all default parameters.

In this default setup CoSMic predicts ICU-cases during the first wave of the COVID-19 pandemic
in the four federal German states "Schleswig-Hollstein", "Hamburg", "Niedersachsen" and "Bremen"
as a kind of regression check.

For an extensive documentation of the model's available parameters please see
https://github.com/hpcralf/CoSMic/blob/main/aux/CoSMic-Parameter-Setup_and_Usage.html

### Remark
As already mentioned above, this demonstration executes CoSMic with default parameters so it is **not
meant to produce quantitatively correct results**. This is due to the default population size of 0.1%
of the German population. It was  found during CoSMic's development that correct prediction of ICU
capacitiy coverage during the first and second wave of the  COVID-19 pandemic for a country like
Germany with ~83 million inhabitants at least requires a 20% population sample. Smaler samples were
found  to run out of healthy individuals in the middle of the second wave. 

First things first. Since we want to be better save than sorry, first of all  we clean the environment.
```{r}
rm(list=ls())
```

Then we load CoSMic. For CoSMic installation instructions please see https://github.com/hpcralf/CoSMic.
```{r}
library(CoSMic)
```