<!--
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%      ___      __         _      
%%     / __\___ / _\  /\/\ (_) ___ 
%%    / /  / _ \\ \  /    \| |/ __|
%%   / /__| (_) |\ \/ /\/\ \ | (__ 
%%   \____/\___/\__/\/    \/_|\___|
%%
%%  COVID-19 Spatial Microsimulation  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Authors:      Christian Dudel
%%               Matthias Rosenbaum-Feldbruegge
%%               Sebastian Kluesener
%%               Ralf Schneider
%%
%% Contact:      dudel@demogr.mpg.de
%%               sebastian.kluesener@bib.bund.de
%%
%%==============================================================================
%%
%% Copyright (C) 2021 Federal Institute for Population Research (BIB),
%%                    The Max Planck Institute for Demographic Research (MPIDR),
%%                    High Performance Computing Center Stuttgart (HLRS)
%% 
%% This program is free software: you can redistribute it and/or modify
%% it under the terms of the GNU General Public License as published by
%% the Free Software Foundation, either version 3 of the License, or
%% (at your option) any later version.
%% 
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%% GNU General Public License for more details.
%% 
%% You should have received a copy of the GNU General Public License
%% along with this program.  If not, see <http://www.gnu.org/licenses/>.
%%==============================================================================
-->
---
title: "CoSMic - Basic usage"
author: "Ralf Schneider <schneider@hlrs.de>"
output:
  html_document:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{CoSMic - Basic usage}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction
In this vignette we demonstrate the usage of the CoSMic function stack to execute
**CO**VID-19 **S**patial **Mic**rosimulations to predict coverage of intensive care units.

At this place we demonstrate the execution of CoSMic with minimal paramter setup efforts
so keeping basically all default parameters.

In this default setup CoSMic predicts ICU-cases during the first wave of the COVID-19 pandemic
in the four federal German states "Schleswig-Hollstein", "Hamburg", "Niedersachsen" and "Bremen"
as a kind of regression check.

For an extensive documentation of the model's available parameters please see
https://github.com/hpcralf/CoSMic/blob/main/aux/CoSMic-Parameter-Setup_and_Usage.html

### Remark
As already mentioned above, this demonstration executes CoSMic with default parameters so it is **not
meant to produce quantitatively correct results**. This is due to the default population size of 0.1%
of the German population. It was  found during CoSMic's development that correct prediction of ICU
capacitiy coverage during the first and second wave of the  COVID-19 pandemic for a country like
Germany with ~83 million inhabitants at least requires a 20% population sample. Smaler samples were
found  to produce siginficantly altered infection dynamics starting in the middle of the second wave. 


# Initialization
First things first. Since we want to be better save than sorry, first of all  we clean the environment.
```{r}
rm(list=ls())
```

Then we load CoSMic. For CoSMic installation instructions please see https://github.com/hpcralf/CoSMic.
```{r}
library(CoSMic)
```

## Setup of Execution Parameters


```{r warning=FALSE}
exec.params <- set.exec.params()
```

## Choose sample size
```{r}
set.pspace(param  = "sam_size", values = 83019)
```

## R0 (basic reproduction number)
```{r}
set.pspace(param  = "R0", values = 3.5)
```

## Duration in ICU
```{r}
set.pspace(param  = "icu_dur", values = 14)
```

## R0 reduction factor
```{r}
data("R0effect.nuts2")
set.pspace(param  = "R0effect",
           values =  R0effect.nuts2)
```

## Weight of infection dynamic
With the following factor the spread of the disease within aregion vs. the
spread to other regions resulting from the connectivity between regions
obtained from commuter flow data is controlled. With setting it to 0.9
90% of internal spreading and 10% of spread to connected regions is selected.
```{r}
set.pspace(param = "w_int", values = 0.9)
```

## Weight of infected cases relocation
With this parameter the relocation of infected cases to othher counties
according to the observerd cases is controlled. With setting it to 0.0
relocation is turned off.
```{r}
set.pspace(param = "w.obs", values = 0.0)
```

## Redistribution method
Either redistribution accross the whole country (0) or only within the
states (1) can be selected.
```{r}
set.pspace(param = "w.obs.by.state", values = 0)
```

## Setup of Static Model Parameters

With \code{set.static.params} we initialize a default parameter list but switch on
all plotting along with full result output.
```{r}
static.params <- set.static.params(
	      pspace=pspace,
	      gplots              = TRUE,
              cplots              = TRUE,
              cplots.states       = TRUE,
              cplots.nuts2        = TRUE,
              results             = "ALL")
```

static.params$results<-"ALL"

## Set input data
```{r}
iol <- load.input()
```

## Generate spatial structure and its population
```{r}
sim.struc <- init.spatial.population( iol , static.params )
```

## Initialize connectivity matrix
```{r}
iol <- init.connectivity( iol , static.params , sim.struc )
```

## Save parameter files
```{r eval=FALSE}
save.exec.params(exec.params)
save.static.params(exec.params, static.params)
save.pspace(exec.params, pspace)
save.input(exec.params, iol)
save.spatial.population(exec.params, sim.struc)
```

## Call CoSMic

### Init parallel execution
```{r eval=FALSE}
init.parallel.execution(exec.params, static.params, NULL)
```

```{r eval=FALSE}
CoSMic (ep        = exec.params,
        sp        = static.params,
        iol       = iol,
	pspace    = pspace,
	sim.struc = sim.struc,
	op = NULL, opt=NULL)
```

## Finalize parallel execution
```{r eval=FALSE}
finalize.parallel.execution(exec.params,cl)
```