<!--
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%      ___      __         _      
%%     / __\___ / _\  /\/\ (_) ___ 
%%    / /  / _ \\ \  /    \| |/ __|
%%   / /__| (_) |\ \/ /\/\ \ | (__ 
%%   \____/\___/\__/\/    \/_|\___|
%%
%%  COVID-19 Spatial Microsimulation  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Authors:      Christian Dudel
%%               Matthias Rosenbaum-Feldbruegge
%%               Sebastian Kluesener
%%               Ralf Schneider
%%
%% Contact:      dudel@demogr.mpg.de
%%               sebastian.kluesener@bib.bund.de
%%
%%==============================================================================
%%
%% Copyright (C) 2021 Federal Institute for Population Research (BIB),
%%                    The Max Planck Institute for Demographic Research (MPIDR),
%%                    High Performance Computing Center Stuttgart (HLRS)
%% 
%% This program is free software: you can redistribute it and/or modify
%% it under the terms of the GNU General Public License as published by
%% the Free Software Foundation, either version 3 of the License, or
%% (at your option) any later version.
%% 
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%% GNU General Public License for more details.
%% 
%% You should have received a copy of the GNU General Public License
%% along with this program.  If not, see <http://www.gnu.org/licenses/>.
%%==============================================================================
-->
---
title: "CoSMic - Basic usage of the Fortran model version"
author: "Ralf Schneider <schneider@hlrs.de>"
output:
  html_document:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{CoSMic - Basic usage of the Fortran model version}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction
In this vignette we demonstrate the usage of the CoSMic Fortran model along with the generation
of the default set of inout files to execute **CO**VID-19 **S**patial **Mic**rosimulations to
predict coverage of intensive care units.

At this place we demonstrate the execution of CoSMic with minimal paramter setup efforts
so keeping basically all default parameters.

In this default setup CoSMic predicts ICU-cases during the first wave of the COVID-19 pandemic
in the sixteen federal German states of Germany as a kind of regression check.

For an extensive documentation of the model's available parameters please see
https://github.com/hpcralf/CoSMic/blob/main/aux/CoSMic-Parameter-Setup_and_Usage.html

### Remark
As already mentioned above, this demonstration executes CoSMic with default parameters so it is **not
meant to produce quantitatively correct results**. This is due to the default population size of 0.1%
of the German population. It was  found during CoSMic's development that correct prediction of ICU
capacitiy coverage during the first and second wave of the  COVID-19 pandemic for a country like
Germany with ~83 million inhabitants at least requires a 20% population sample. Smaler samples were
found  to produce siginficantly altered infection dynamics starting in the middle of the second wave. 

# Initialization

To be consistent with the R-version of the model the R function `convert.Rp.to.Fp()` is provided that
converts the R-version's parameter sets as well as the input-files into parameter- and input-files which
are compatible with the Fortran-version.

First things first. Since we want to be better save than sorry, first of all  we clean the environment.
```{r}
rm(list=ls())
```

Then we load CoSMic. For CoSMic installation instructions please see https://github.com/hpcralf/CoSMic.
```{r}
library(CoSMic)
```

## Preparation of parameter- and input files

### Setup of Execution Parameters

In this basic usage scenario we set up the default execution parameters.
```{r}
exec.params <- set.exec.params()
```

### Setup of the parameter space

In the pspace parameter list, model parameters are stored that are potentially subject to variations.
In this basic usage scenario the default list provided by the CoSMic package can be used. To avoid
naming confilcts the param.space provided by the CoSMic package has to be renamed to `pspace`.

```{r}
data("param.space")
pspace <- param.space
```

### Setup of Static Model Parameters

With `set.static.params()` we initialize a default parameter list.
```{r}
static.params <- set.static.params(pspace=pspace)
```

### Set input data
As model input we use the datasets provided by the CoSMic package. they can be initalized by a
call to `load.input()`.
```{r}
iol <- load.input()
```

### Generate spatial structure, its population and connectivity
The missing steps are to initialize the population with its spatial structure and the connectivity
betwenn the spatial compartments.
```{r}
sim.struc <- init.spatial.population( iol , static.params )
iol <- init.connectivity( iol , static.params , sim.struc )
```

### Save parameter files
To save the initialized parameter sets for later usage we save them in RDS format. The files will
be stored in the output directory, genearted during the call to `set.exec.params()`.
```{r eval=FALSE}
save.exec.params(exec.params)
save.static.params(exec.params, static.params)
save.pspace(exec.params, pspace)
save.input(exec.params, iol)
save.spatial.population(exec.params, sim.struc)
```

### Generate parameter- and input-files

```{r eval=FALSE}
convert.Rp.to.Fp(filename.sp = "static_parameters.dat",
                 sp          = static.params,
                 filename.ep = "exec_parameters.dat",
		 ep          = exec.params,
                 iol         = iol,
		 pspace      = pspace,
                 outpath     = exec.params$output.dir)
```

## Execute Fortran-version

Provided the CoSMic binary is in the execution path of your environment the
Fortran-version can be executed from within the directory in which the input
files for the execution and static parameters reside.
```{bash eval=FALSE}
export OMP_NUM_THREADS=4
mpirun -n 1 CoSMic
```
